
---
title: "ARMA-GARCH Modeling of SSE Composite Index Returns"
author: "Wei Yuan"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
  pdf:
    toc: true
    number-sections: true
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
source("../src/00_setup.R")
load_packages()
set_report_options()

source("../src/01_data_prep.R")
source("../src/02_split.R")
source("../src/03_stats_tests.R")
source("../src/04_model_specs.R")
source("../src/05_fit_select.R")
source("../src/06_diagnostics.R")
source("../src/07_rolling_forecast.R")
source("../src/08_plots.R")
```

```{r pipeline_objects, include=FALSE}
prep <- load_and_prepare_sse("../data/SSE_Jan2008_Oct2025.xlsx")

split_obj <- split_series(prep$sse_xts, train_frac = 0.70)

desc_obj <- compute_descriptive_stats(split_obj$sse_train)
adf_res <- run_adf_test(split_obj$sse_train)
acf_pacf_obj <- get_acf_pacf(split_obj$sse_train)

combos <- get_candidate_models()

tournament <- run_model_tournament(
  combos = combos,
  sse_train = split_obj$sse_train,
  sse_xts = prep$sse_xts,
  sse_test = split_obj$sse_test,
  n_train = split_obj$n_train,
  n_total = split_obj$n_total
)

best_model_name <- tournament$best_model_name
best_fit_obj <- tournament$best_fit_obj

diag_obj <- run_model_diagnostics(best_fit_obj)

roll_obj <- run_rolling_forecast(
  best_model_name = best_model_name,
  combos = combos,
  sse_xts = prep$sse_xts,
  n_train = split_obj$n_train
)
```

# Introduction

This report studies the daily return dynamics and volatility of the Shanghai Stock Exchange (SSE) Composite Index using a joint ARMA-GARCH framework. The objective is to identify a model that is both statistically defensible and practically useful for out-of-sample risk forecasting.

The analysis is organized as a reproducible research pipeline. Data preparation, model fitting, diagnostics, and forecasting logic are implemented in reusable source files, while this report focuses on presenting results and interpretation.

# Data Preparation

The raw input consists of daily SSE index price data. The cleaning stage standardizes date formats, converts prices to numeric values, removes problematic rows, sorts the data chronologically, and computes daily log returns.

```{r data_preparation}
cat(sprintf("Rows processed: %d\n", nrow(prep$df)))
cat(sprintf("Date range: %s to %s\n", min(index(prep$sse_xts)), max(index(prep$sse_xts))))
cat(sprintf("Total return observations: %d\n\n", NROW(prep$sse_xts)))

head(prep$df)
```

The return series used in the remainder of the analysis is stored as an `xts` object.

```{r preview_return_series}
head(prep$sse_xts)
```

# Train-Test Split

To evaluate predictive performance out of sample, the return series is split into a training set and a testing set. The training sample is used for model estimation and selection, while the testing sample is reserved for forecast evaluation.

```{r train_test_split}
plot_train_test_split(prep$sse_xts, split_obj$sse_train)

cat(sprintf("Training observations: %d\n", NROW(split_obj$sse_train)))
cat(sprintf("Testing observations: %d\n", NROW(split_obj$sse_test)))
cat(sprintf("Training fraction: %.0f%%\n", 100 * split_obj$n_train / split_obj$n_total))
```

# Descriptive Statistics

Before fitting ARMA-GARCH models, it is useful to inspect the basic empirical properties of the training data.

```{r descriptive_stats}
knitr::kable(
  desc_obj$stats,
  caption = "Descriptive Statistics for SSE Training Returns"
)

cat("Excess kurtosis:", desc_obj$kurt_val, "\n")
```

The return distribution is expected to exhibit fat tails, which motivates the use of Student-t innovations in the volatility models.

# Stationarity Check

A return series used in ARMA-GARCH modeling should be stationary in mean. This is assessed using the Augmented Dickey-Fuller test.

```{r stationarity_test}
print(adf_res)
```

A strongly negative test statistic with a small p-value supports the assumption that the return series is stationary.

# Preliminary Mean Identification

Autocorrelation and partial autocorrelation plots are used as a preliminary guide for plausible mean specifications. The final mean-variance specification is still determined jointly through the model tournament.

```{r model_identification}
plot_acf_pacf(split_obj$sse_train)
```

These plots help motivate low-order ARMA candidates rather than providing a definitive model selection rule.

# Candidate Model Set

The project compares a compact set of joint mean-variance specifications:

* AR(1)-GARCH(1,1)
* MA(1)-GJR-GARCH(1,1)
* ARMA(1,1)-EGARCH(1,1)
* AR(1)-GARCH-in-Mean

All models are estimated with Student-t innovations to account for non-Gaussian tails.

```{r candidate_models}
str(combos)
```

# Joint Model Tournament

Each candidate model is fit on the training set. Model comparison uses:

* **BIC** for in-sample parsimony
* **RMSE** on the test set for out-of-sample predictive performance

The main decision criterion is out-of-sample RMSE.

```{r grand_tournament}
knitr::kable(
  tournament$results,
  caption = "Joint Model Selection Results"
)

cat("Selected model based on lowest test RMSE:", best_model_name, "\n")
```

This stage is the core of the research pipeline: the mean and variance equations are selected jointly rather than in isolation.

# Final Estimated Model

The winning specification is re-used as the main fitted model for interpretation and diagnostic checking.

```{r final_estimation}
show(best_fit_obj)
```

The output above provides the estimated parameters of the selected ARMA-GARCH specification.

# Residual Diagnostics

A fitted model should not only forecast well, but also leave residuals that are approximately free of serial dependence and leftover ARCH structure.

Two main checks are used:

* **Ljung-Box test** on standardized residuals
* **ARCH-LM test** on standardized residuals

```{r full_model_diagnostics}
print(diag_obj$lb_test)
print(diag_obj$arch_test)
```

Autocorrelation plots of standardized residuals and squared standardized residuals provide a visual complement to the formal tests.

```{r diagnostic_plots}
plot_diagnostics(diag_obj$std_resid)
```

A satisfactory model should show little structure remaining in either residual series.

# Rolling Volatility Forecast

After selecting the best model, the next step is to evaluate it in a rolling forecasting setting. One-step-ahead recursive forecasts are generated using periodic re-estimation.

```{r rolling_forecast}
head(roll_obj$forecast_df)
```

This produces a time series of predicted conditional volatility over the testing period.

# Forecast vs Realized Returns

The plot below compares predicted conditional volatility with the realized return series in the test sample.

```{r perf_plot}
plot_forecast_vs_returns(
  forecast_df = roll_obj$forecast_df,
  sse_xts = prep$sse_xts,
  sse_test = split_obj$sse_test,
  n_train = split_obj$n_train,
  n_total = split_obj$n_total
)
```

The conditional volatility forecast is not intended to predict the sign of returns directly. Instead, it captures the changing scale of risk through time.

# Discussion

Several features of the SSE return series make ARMA-GARCH modeling appropriate:

* returns are stationary after transformation from prices
* fat tails motivate Student-t innovations
* volatility clustering motivates conditional heteroskedastic models
* asymmetric or nonlinear variance dynamics may justify alternatives such as GJR-GARCH or EGARCH

The model tournament balances econometric structure with predictive usefulness by combining in-sample and out-of-sample criteria.

# Conclusion

This project builds a modular research pipeline for modeling SSE Composite Index returns using ARMA-GARCH family models. The workflow includes:

* robust data cleaning and return construction
* train-test evaluation design
* descriptive and stationarity analysis
* joint ARMA-GARCH model selection
* residual diagnostics
* rolling one-step-ahead volatility forecasting

The refactored structure separates reusable research code from reporting logic, making the analysis easier to maintain, extend, and present.

# Reproducibility

The full project is organized so that:

* `src/` contains reusable analysis functions
* `scripts/run_pipeline.R` runs the full workflow without rendering the report
* this QMD serves as the presentation layer for results and interpretation

To rerun the report from the repository root:

```{r eval=FALSE}
quarto::quarto_render("reports/arma_garch_report.qmd")
```

