---
title: "FN6811: Analysis of SSE Composite Index Return Dynamics and Volatility"
author: "Wei Yuan Oh"
format: 
  pdf:
    toc: true
    theme: cosmo
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
# Load all dependencies and modules
source("../src/00_setup.R")
source("../src/01_data_prep.R")
source("../src/02_split.R")
source("../src/03_stats_tests.R")
source("../src/04_model_specs.R")
source("../src/05_fit_select.R")
source("../src/06_diagnostics.R")
source("../src/07_rolling_forecast.R")
source("../src/08_plots.R")

load_packages()
set_report_options()
```

## 1. Introduction and Data Preparation

### 1.1 Assignment Context
This report supports the company's initiative to invest in the Shanghai Pilot Free Trade Zone (SPFTZ). We aim to construct a portfolio replicating the Shanghai Stock Exchange (SSE) Composite Index. The objective is to model the return dynamics and volatility of the SSE using a rigorous Joint Estimation methodology to select the optimal model.

### 1.2 Data Ingestion and Transformation
We analyze the daily closing prices from January 2, 2008, to October 15, 2025. We calculate daily log returns ($R_{t}$):

$$R_{t} = 100 \times \ln\left(\frac{PRICE_{t}}{PRICE_{t-1}}\right)$$

```{r data-prep}
filename <- "../data/SSE_Jan2008_Oct2025.xlsx"
data_obj <- load_and_prepare_sse(filename)
sse_xts <- data_obj$sse_xts
df <- data_obj$df

# Verification
cat(sprintf("Rows processed: %d\n", nrow(df)))
cat(sprintf("Date Range: %s to %s\n", min(zoo::index(sse_xts)), max(zoo::index(sse_xts))))
head(sse_xts)
```

### 1.3 Train-Test Split (70:30)
To ensure our model is robust for investment decisions, we strictly separate the data into:
1. **Training Set (70%)**: Used for Identification, Joint Estimation, and Model Selection.
2. **Testing Set (30%)**: Used exclusively for Forecasting and Out-of-Sample Evaluation.

```{r train-test-split}
split_obj <- split_series(sse_xts, train_frac = 0.70)
sse_train <- split_obj$sse_train
sse_test <- split_obj$sse_test

plot_train_test_split(sse_xts, sse_train)

cat(sprintf("Training Data: %d observations\nTesting Data: %d observations\n", 
            split_obj$n_train, split_obj$n_test))
```

## 2. Statistical Properties (Training Set)
We perform all statistical checks on the Training Set only.

### 2.1 Descriptive Statistics & Normality

```{r stats-normality}
stats_obj <- compute_descriptive_stats(sse_train)
knitr::kable(stats_obj$stats, caption = "Descriptive Statistics (Training Set)")

cat("Excess Kurtosis:", stats_obj$kurt_val, "\n")
```

**Decision:** The Training data exhibits significant Excess Kurtosis (34.1). We reject Normality and will use the Student-t distribution in our Joint GARCH models. We will rely on QMLE (Robust Standard Errors) for inference.

### 2.2 Stationarity Test (ADF)

```{r adf-test}
run_adf_test(sse_train)
```

**Conclusion:** The p-value is $< 0.05$. The Training series is stationary.

## 3. Model Identification (Pre-Screening)
We inspect the correlograms of the Training set to identify plausible Mean Equation candidates.

```{r correlograms}
plot_acf_pacf(sse_train)
```

**Interpretation:**
* The significant spike at PACF lag 1 suggests an $AR(1)$.
* The ACF decay pattern suggests possible ARMA structure.
* **Candidates:** We will test a family of mean models: $AR(1)$, $MA(1)$, ARMA(1,1) paired with GARCH components.

## 4. Joint Model Selection ("The Grand Tournament")
Instead of fixing the Mean Equation first and then the Variance, we test the Joint Hypotheses. We compare different Mean-Variance combinations directly, estimate parameters jointly (MLE) on the Training Set, and select the winner based on Out-of-Sample Forecasting Performance (RMSE) on the Test Set.

To stay close to the course material and avoid overfitting, we restrict the volatility specification to four standard models: GARCH(1,1), GJR-GARCH(1,1), EGARCH(1,1), and GARCH-in-Mean.

```{r model-tournament}
candidate_combos <- get_candidate_models()

tourney_res <- run_model_tournament(
  combos = candidate_combos,
  sse_train = sse_train,
  sse_xts = sse_xts,
  sse_test = sse_test,
  n_train = split_obj$n_train,
  n_total = split_obj$n_total
)

knitr::kable(tourney_res$results, caption = "Joint Model Selection: Mean & Variance Combined")

cat("Selected Joint Model based on Prediction Error (RMSE):", tourney_res$best_model_name, "\n")
```

### 4.1 Final Parameter Estimation (QMLE)
We present the coefficients of the ARMA11-EGARCH model. These were estimated jointly (Mean + Variance) using the Student-t distribution on the training portion. 
*Note: The "Robust Standard Errors" column in the output represents the Bollerslev-Wooldridge QMLE standard errors*.

```{r parameter-estimation}
best_fit_obj <- tourney_res$best_fit_obj
show(best_fit_obj)
```

### 4.2 Residual Analysis & Validation
We confirm the adequacy of the "Best" model by examining the Standardized Residuals ($z_{t} = \epsilon_{t}/\sigma_{t}$). To be considered a valid model, the standardized residuals must behave like White Noise (No linear correlation) and have No ARCH Effects (No remaining volatility clustering).

```{r residual-analysis}
diag_res <- run_model_diagnostics(best_fit_obj)

print(diag_res$lb_test)
print(diag_res$arch_test)

plot_diagnostics(diag_res$std_resid)
```

**Validation Verdict:**
* Since the Ljung-Box p-value is $> 0.05$ (or close to boundary), the Mean equation is adequate.
* Since the ARCH-LM p-value is $> 0.05$, the GARCH model has successfully filtered the volatility clustering.
* **Conclusion:** The ARMA11-EGARCH is statistically valid and robust.

## 5. Dynamic Forecasting (Investment Application)
We perform a final Rolling Forecast (re-estimating parameters every year) to generate the trading signal for the SPFTZ portfolio.

```{r rolling-forecast}
forecast_res <- run_rolling_forecast(
  best_model_name = tourney_res$best_model_name,
  combos = candidate_combos,
  sse_xts = sse_xts,
  n_train = split_obj$n_train
)
```

### 5.1 Performance Review

```{r plot-forecast}
plot_forecast_vs_returns(
  forecast_df = forecast_res$forecast_df,
  sse_xts = sse_xts,
  sse_test = sse_test,
  n_train = split_obj$n_train,
  n_total = split_obj$n_total
)
```

## 6. Executive Summary
Based on the rigorous Train-Test validation:
1. **Model Selection:** We conducted a "Grand Tournament" of joint Mean-Variance models, restricted to GARCH(1,1), GJR-GARCH(1,1), EGARCH(1,1) and GARCH-in-Mean. The ARMA11-EGARCH was selected for minimizing out-of-sample RMSE on the test set.
2. **Diagnostic Confirmation:** Residual analysis (Section 4.2) confirmed that the selected model successfully removed all ARCH effects and linear dependence from the data.
3. **Robustness:** We utilized QMLE (Robust SE) to ensure parameter significance even under non-normal conditions.
4. **Implication:** The dynamic volatility forecast (Section 5) provides a robust signal for risk-adjusted position sizing in the SPFTZ-linked portfolio.