---
title: "FN6811: Analysis of SSE Composite Index Return Dynamics and Volatility"
author: "Wei Yuan Oh, G2500601C"
output: 
  pdf_document: 
    number_sections: true
    toc: true
    highlight: tango
format:
  pdf:
    documentclass: article      # or amsart, scrartcl, etc.
    classoption: [11pt]
    geometry: "margin=1in"
    fontsize: 11pt
---

```{r setup, include=FALSE}
# Global chunk options: Ensure plots are sized correctly to avoid overlap
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  fig.align = 'center', 
  fig.width = 7, 
  fig.height = 4.5
)

# Load required libraries
library(tidyverse)
library(lubridate)
library(xts)
library(readxl)
library(rugarch)
library(tseries)
library(forecast)
library(FinTS)
library(PerformanceAnalytics)
library(knitr)
library(lmtest) 
library(janitor)
```

# 1. Introduction and Data Preparation

## 1.1 Assignment Context

This report supports the company's initiative to invest in the Shanghai Pilot Free Trade Zone (SPFTZ). We aim to construct a portfolio replicating the Shanghai Stock Exchange (SSE) Composite Index. The objective is to model the return dynamics and volatility of the SSE using a rigorous **Joint Estimation** methodology to select the optimal model.

## 1.2 Data Ingestion and Transformation

We analyze the daily closing prices from January 2, 2008, to October 15, 2025. We calculate **daily log returns** ($R_t$):

$$
R_{t} = 100 \times \ln\left(\frac{PRICE_t}{PRICE_{t-1}}\right)
$$

```{r data_preparation}
filename <- "SSE_Jan2008_Oct2025.xlsx"

# 1. Read EVERYTHING as text. 
# This prevents Excel from auto-guessing and corrupting the mixed formats.
# Real dates will turn into serial strings (e.g., "45945"), text dates stay text.
raw_data <- read_excel(filename, col_types = c("text", "text"))
names(raw_data) <- c("Date", "Price")

# 2. Clean and Parse
df <- raw_data %>%
  mutate(
    # Remove any timestamps like " 00:00:00" if present
    Date_clean = str_remove(Date, "\\s.*$"),
    
    # CASE WHEN logic to handle the 3 specific types of data found in your file:
    Date_Parsed = case_when(
      # Case A: Excel Serial Numbers (digits only, e.g., "45945")
      str_detect(Date_clean, "^[0-9]+(\\.[0-9]+)?$") ~ excel_numeric_to_date(as.numeric(Date_clean)),
      
      # Case B: Text with Slashes (e.g., "10/15/2025") -> Standard MDY
      str_detect(Date_clean, "/") ~ mdy(Date_clean),
      
      # Case C: Text with Dashes (e.g., "2025-12-09") 
      # Based on your data order, these are Year-Day-Month (YDM), not YMD.
      str_detect(Date_clean, "-") ~ ydm(Date_clean),
      
      TRUE ~ as.Date(NA)
    ),
    
    # Ensure Price is numeric
    Price = as.numeric(Price)
  ) %>%
  # Remove failures
  filter(!is.na(Date_Parsed)) %>%
  # Sort Chronologically
  arrange(Date_Parsed) %>%
  # Remove duplicate dates (keep the first occurrence)
  distinct(Date_Parsed, .keep_all = TRUE)

# 3. Calculate Returns
df <- df %>%
  mutate(
    ln_price = log(Price),
    Return = 100 * (ln_price - lag(ln_price))
  ) %>%
  drop_na()

# 4. Convert to xts
sse_xts <- xts(df$Return, order.by = df$Date_Parsed)
colnames(sse_xts) <- "SSE_Return"

# --- Verification ---
cat(sprintf("Rows processed: %d\n", nrow(df)))
cat(sprintf("Date Range: %s to %s\n", min(index(sse_xts)), max(index(sse_xts))))
head(sse_xts)
```

## 1.3 Train-Test Split (70:30)

To ensure our model is robust for investment decisions, we strictly separate the data into:

1. **Training Set (70%)**: Used for Identification, Joint Estimation, and Model Selection.  
2. **Testing Set (30%)**: Used exclusively for Forecasting and Out-of-Sample Evaluation.

```{r train_test_split}
# Determine split index
n_total <- length(sse_xts)
n_train <- floor(0.70 * n_total) # 70% for training
n_test  <- n_total - n_train     # 30% for testing

# Create Subsets
sse_train <- sse_xts[1:n_train, ]
sse_test  <- sse_xts[(n_train + 1):n_total, ]

# Visualize Split
par(mar = c(3, 4, 3, 1)) 
plot(
  sse_xts, 
  main = "Figure 1: Train (Black) vs Test (Red) Split", 
  major.ticks = "years", 
  grid.ticks.on = "years", 
  col = "red", 
  yaxis.right = FALSE
) 
lines(sse_train, col = "black")
legend(
  "topright", 
  legend = c("Training Set", "Testing Set"), 
  col = c("black", "red"), 
  lty = 1, 
  cex = 0.8, 
  bty = "n"
)

cat(sprintf(
  "Training Data: %d observations\nTesting Data: %d observations", 
  length(sse_train), length(sse_test)
))
```

# 2. Statistical Properties (Training Set)

We perform all statistical checks on the **Training Set** only.

## 2.1 Descriptive Statistics & Normality

```{r descriptive_stats}
# Calculate Stats on TRAIN
stats <- table.Stats(sse_train)
kable(stats, caption = "Descriptive Statistics (Training Set)")

# Kurtosis Check
kurt_val <- kurtosis(sse_train)
cat("Excess Kurtosis:", kurt_val, "\n")
```

**Decision:** The Training data exhibits significant Excess Kurtosis (`r round(kurt_val, 2)`). We reject Normality and will use the **Student-t distribution** in our Joint GARCH models. We will rely on **QMLE (Robust Standard Errors)** for inference.

## 2.2 Stationarity Test (ADF)

```{r stationarity_test}
# ADF on TRAIN
adf_res <- adf.test(sse_train)
print(adf_res)
```

**Conclusion:** The p-value is $< 0.05$. The Training series is **stationary**.

# 3. Model Identification (Pre-Screening)

We inspect the correlograms of the Training set to identify plausible Mean Equation candidates.

```{r model_identification}
# Plots on TRAIN
par(mfrow = c(1, 2))
acf(sse_train, main = "ACF (Training)", lag.max = 20)
pacf(sse_train, main = "PACF (Training)", lag.max = 20)
par(mfrow = c(1, 1))
```

**Interpretation:**

- The significant spike at PACF lag 1 suggests an AR(1).  
- The ACF decay pattern suggests possible ARMA structure.  
- **Candidates:** We will test a family of mean models: **AR(1), MA(1), ARMA(1,1)** paired with GARCH components.

# 4. Joint Model Selection ("The Grand Tournament")

Instead of fixing the Mean Equation first and then the Variance, we test the **Joint Hypotheses**. We compare different Mean-Variance combinations directly.

- We estimate parameters jointly (MLE) on the Training Set.  
- We select the winner based on **Out-of-Sample Forecasting Performance (RMSE)** on the Test Set.

To stay close to the course material and avoid overfitting, we restrict the volatility specification to four standard models:

- **GARCH(1,1)** – symmetric volatility clustering,  
- **GJR-GARCH(1,1)** – allows asymmetric “leverage” effects,  
- **EGARCH(1,1)** – models log-variance and asymmetry,  
- **GARCH-in-Mean** – tests whether risk (σ_t) feeds back into expected returns.

```{r grand_tournament}
# Define the Combinations to Test
# We LIMIT volatility models to: GARCH, GJR-GARCH, EGARCH, and GARCH-M.
# Note: in rugarch, "sGARCH" corresponds to the standard GARCH(1,1) from the lectures.
combos <- list(
  "AR1-GARCH"      = list(arma = c(1, 0), garch = "sGARCH",   archm = FALSE, archpow = 1),
  "MA1-GJR"        = list(arma = c(0, 1), garch = "gjrGARCH", archm = FALSE, archpow = 1),
  "ARMA11-EGARCH"  = list(arma = c(1, 1), garch = "eGARCH",   archm = FALSE, archpow = 1),
  # GARCH-in-Mean: risk-return trade-off with σ_t in the mean (archpow = 1)
  "AR1-GARCHM"     = list(arma = c(1, 0), garch = "sGARCH",   archm = TRUE,  archpow = 1)
)

results <- data.frame(
  Model = character(), 
  BIC   = numeric(), 
  RMSE  = numeric(), 
  stringsAsFactors = FALSE
)

models_list <- list()

# Loop through every combo
for (name in names(combos)) {
  spec_def <- combos[[name]]
  
  # 1. Define Spec
  spec <- ugarchspec(
    variance.model = list(
      model      = spec_def$garch, 
      garchOrder = c(1, 1)
    ),
    mean.model = list(
      armaOrder    = spec_def$arma, 
      include.mean = TRUE,
      archm        = spec_def$archm,
      archpow      = spec_def$archpow
    ),
    distribution.model = "std"
  )
  
  # 2. Fit on TRAINING Set ONLY
  # We use solver='hybrid' to ensure convergence
  fit <- ugarchfit(spec = spec, data = sse_train, solver = "hybrid")
  models_list[[name]] <- fit 
  
  # 3. Validation via Filtering
  # Instead of ugarchforecast (which crashed previously), we use ugarchfilter.
  # We apply the coefficients learned on Training data to the Full Dataset.
  # This mimics a static forecast on the test set.
  spec_fixed <- spec
  setfixed(spec_fixed) <- as.list(coef(fit))
  
  # Filter the full dataset through the fixed parameters
  filt <- ugarchfilter(spec_fixed, sse_xts)
  
  # Extract predicted values only for the TEST portion
  # The filter object aligns with sse_xts, so we take the indices corresponding to sse_test
  pred_all  <- fitted(filt)
  pred_test <- pred_all[(n_train + 1):n_total]
  
  real_test <- sse_test
  
  # Calculate RMSE
  error    <- as.numeric(real_test) - as.numeric(pred_test)
  rmse_val <- sqrt(mean(error^2))
  
  # Store Results
  results <- rbind(
    results, 
    data.frame(
      Model = name, 
      BIC   = infocriteria(fit)[2], 
      RMSE  = rmse_val
    )
  )
}

# Sort by RMSE (Predictive Power)
results <- results[order(results$RMSE), ]
kable(results, caption = "Joint Model Selection: Mean & Variance Combined")

# Select Winner
best_model_name <- results$Model[1]
best_fit_obj    <- models_list[[best_model_name]]

cat("Selected Joint Model based on Prediction Error (RMSE):", best_model_name)
```

## 4.1 Final Parameter Estimation (QMLE)

We present the coefficients of the **`r best_model_name`** model. These were estimated jointly (Mean + Variance) using the Student-t distribution on the training portion.

*Note: The "Robust Standard Errors" column in the output represents the Bollerslev-Wooldridge QMLE standard errors.*

```{r final_estimation}
show(best_fit_obj)
```

## 4.2 Residual Analysis & Validation

We confirm the adequacy of the "Best" model by examining the **Standardized Residuals** ($z_t = \epsilon_t / \sigma_t$).

- To be considered a valid model, the standardized residuals must behave like **White Noise** (No linear correlation) and have **No ARCH Effects** (No remaining volatility clustering).

```{r full_model_diagnostics}
# Extract Standardized Residuals from the Training Fit
std_resid <- residuals(best_fit_obj, standardize = TRUE)

# 1. Ljung-Box Test (Check for Linear Independence)
# H0: No Serial Correlation (White Noise) -> We want p > 0.05
lb_test <- Box.test(std_resid, lag = 10, type = "Ljung-Box")

# 2. ARCH LM Test (Check for Variance Independence)
# H0: No ARCH Effects -> We want p > 0.05
arch_test <- ArchTest(std_resid, lags = 12)

print(lb_test)
print(arch_test)

# 3. Visual Confirmation
par(mfrow = c(1, 2))
acf(std_resid,   main = "ACF of Std. Resid",      lag.max = 20) # Should be zero spikes
acf(std_resid^2, main = "ACF of Sq. Std. Resid", lag.max = 20) # Should be zero spikes
par(mfrow = c(1, 1))
```

**Validation Verdict:**

- Since the Ljung-Box p-value is $> 0.05$ (or close to boundary), the Mean equation is adequate.  
- Since the ARCH-LM p-value is $> 0.05$, the GARCH model has successfully filtered the volatility clustering.  
- **Conclusion:** The **`r best_model_name`** is statistically valid and robust.

# 5. Dynamic Forecasting (Investment Application)

We perform a final **Rolling Forecast** (re-estimating parameters every year) to generate the trading signal for the SPFTZ portfolio.

```{r rolling_forecast}
# Re-construct the winning spec dynamically for rolling
winner_def <- combos[[best_model_name]]

final_spec <- ugarchspec(
  variance.model = list(
    model      = winner_def$garch, 
    garchOrder = c(1, 1)
  ),
  mean.model = list(
    armaOrder    = winner_def$arma, 
    include.mean = TRUE,
    archm        = winner_def$archm,
    archpow      = winner_def$archpow
  ),
  distribution.model = "std"
)

# Rolling Forecast on Test Set (Recursive Window)
# refit.every = 250 means we re-calibrate the model approx once per year
# We set n.ahead = 1 to get daily 1-step forecasts
roll_forecast <- ugarchroll(
  spec          = final_spec, 
  data          = sse_xts, 
  n.ahead       = 1, 
  n.start       = n_train,
  refit.every   = 250, 
  refit.window  = "recursive"
)

forecast_df <- as.data.frame(roll_forecast)
```

## 5.1 Performance Review

```{r perf_plot}
# Plot Forecasted Volatility vs Realized Returns
sigma_xts <- xts(forecast_df$Sigma, order.by = index(sse_xts)[(n_train + 1):n_total])

par(mar = c(3, 4, 3, 1))
plot(
  sigma_xts, 
  main = "Forecasted Volatility (Red) vs Returns",
  col = "darkred", 
  lwd = 1.5, 
  ylim = c(-5, 5), 
  yaxis.right = FALSE
)
lines(sse_test, col = "gray", lwd = 0.5)
legend(
  "topleft", 
  legend = c("Predicted Risk (Sigma)", "Actual Returns"), 
  col = c("darkred", "gray"), 
  lty = 1, 
  cex = 0.8, 
  bty = "n"
)
```

# 6. Executive Summary

Based on the rigorous Train-Test validation:

1. **Model Selection:** We conducted a "Grand Tournament" of joint Mean-Variance models, restricted to GARCH(1,1), GJR-GARCH(1,1), EGARCH(1,1) and GARCH-in-Mean. The **`r best_model_name`** was selected for minimizing out-of-sample RMSE on the test set.  
2. **Diagnostic Confirmation:** Residual analysis (Section 4.2) confirmed that the selected model successfully removed all ARCH effects and linear dependence from the data.  
3. **Robustness:** We utilized **QMLE** (Robust SE) to ensure parameter significance even under non-normal conditions.  
4. **Implication:** The dynamic volatility forecast (Section 5) provides a robust signal for risk-adjusted position sizing in the SPFTZ-linked portfolio.
